---
title: "Assignment_5"
author: "Juhi Bhattacharya"
date: "2024-09-12"
output: html_document
---
Use rnorm to sample from a Normal distribution with mean = 100 and standard deviation 15. 

But I want you to systematically vary sample size as follows (variable n): 1, 5, 10, 100, 1000, 10000

And also systematically vary the experiment size (variable i): 10, 100, 1000, 10000, 100000

For each sample, compute the mean, standard deviation and plot the resulting distribution. 

Note down your remarks at the end on what happens with increase in i and n.

```{r}
sample_sizes <- c(1, 5, 10, 100, 1000, 10000)
experiment_sizes <- c(10, 100, 1000, 10000, 100000)

means <- 0 
std <- 0
j <- 0

for (x in 1:5) {
  j <- experiment_sizes[x]
  for (i in 1:j) {
    IQ.n <- round(rnorm(n=sample_sizes, mean = 100, sd = 15))
    means[i] <- mean(IQ.n)
    std[i] <- sd(IQ.n)
}
hist_means <- hist(means)
plot(hist_means$mids, hist_means$density, type='l',col='blue',lwd=3 )
  
}
```
# REMARK: Histograms of means of all samples. As we increase n and i it is going to become more narrowed. It is converging towards true mean (almost looks like normmal distribution).

```{r}
means <- 0 #Initializing vector # i =no. of samples/experiment size
for (i in 1:10) {
  IQ.n <- round(rnorm(n=1, mean = 100, sd = 15))
  s[i] <- sqrt(((IQ.n-mean(IQ.n))^2)/length(IQ.n))
  means[i] <- mean(IQ.n)
}
hist_means <- hist(means)
hist_sd <- hist(s)
plot(hist_means$mids, hist_means$density, type='l',col='blue',lwd=3 )
```

```{r}
means <- 0 #Initializing vector # i =no. of samples/experiment size
ssd <- 0
for (i in 1:10) {
  IQ.n <- round(rnorm(n=1, mean = 100, sd = 15))
  s[i] <- ((IQ.n-mean(IQ.n))^2)/length(IQ.n)
  ssd[i] <- sd(IQ.n)
  means[i] <- mean(IQ.n)
}
hist_means <- hist(means)
hist_sd <- hist(s)
hist_ssd <- hist(ssd)
plot(hist_means$mids, hist_means$density, type='l',col='blue',lwd=3 )
```

